Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-06-14 Paper.Rmd
changes in contrasts datasets handling

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}). The table shows that BB1 and SBB1 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating nearly symmetric tail dependence. BB1, being as a two parameter copula,  has both lower and upper tail dependence and depending upon which parameter is greater it shows comparatively greater lower or upper tail dependence. SBB1 is just opposite of BB1 in tail dependency. So, the result was expected, because the contributions came from more or less equally weighted BB1 and SBB1. We also tested that lowest-AIC copula BB1 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

@@

Results indicate that dependence of data is different in some bins than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. But Table \ref{SM-tab_BMR_asym} indicates that tail dependence for the BMR data copula is significantly (95% CI) not different from symmetric normal copula and thus consistent with model selection results.

@@

The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.

>>>>

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}). The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence. We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

@@

Results indicate that dependence of data is different in some bins than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.

@@

The test of indepedence gave $p=0$ <!--`r BivMS_contrasts_67$IndepTestRes` this is e-15 could not compile latex--> (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}). The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was <!--$\num{1e-10}$--> `r round(BivMS_contrasts_67$relLTdep_AICw,3)` <!--`r BivMS_contrasts_67$relLTdep_AICw` this is e-4 could not compile latex--> and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.

### 2018-06-14 SupportingInformation.Rmd
changes in contrasts datasets handling

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\caption[\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs. ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs. ln(mass) contrasts]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs. ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs. ln(mass) contrasts. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption[\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

stat_contrasts_67_ga_numbin3<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin2",numbin = 3)

>>>>

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\caption[\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs. ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs. ln(mass) contrasts taken as their absolute values]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs. ln(mass) of contrasts \textbf{(B)} Copula plot for absolute values of ln(BMR) vs. ln(mass) for all $`r nrow(v_BMR)`$ contrasts. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\caption[\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits(taken as absolute values)]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of absolute values of contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript). \label{fig_raw_cop_AVG_humerus_ltoe}}
\end{center}
\end{figure}

@@

stat_contrasts_67_ga_numbin3<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin3",numbin = 3)
